\section{Conclusion}
\label{conclusion}

\joerg{0.1--25 page: adapt ijcai version.}

\joerg{ijcai text:}

%% \joerg{use some of the following text pars as a brief post-hoc
%%   discussion of related work? ... probably not no space. domaina
%%   nalysis and model checking are briefly mentioned int he intro; Brian
%%   is briefly mentioned in the goal dependencies section. I don't think
%%   we can do better than this.}
%
%% % current intro text
%% %
%% %% We remark that our approach can be viewed as a generalization of
%% %% domain/task analysis (as done \eg\ by \cite{fox:long:jair-98}), and as
%% %% an instance of model checking applied to planning models (related
%% %% \eg\ to \cite{vaquero:etal:keq-13}). Our contribution lies in
%% %% introducing this intermediate problem suited to XAIP as outlined, and
%% %% instantiating that formulation with initial technology showing promise
%% %% in practice.
%
%% Another alternate view of our approach is as a form of domain analysis
%% (actually: task analysis), identifying particular properties of plan
%% space ahead of time. Indeed, various popular task analyses can be cast
%% as instances of our framework. A fact pair $(p,q)$ is mutually
%% exclusive \cite{blum:furst:ai-97} iff $p$-true-at-end entails $\neg
%% q$-true-at-end in the space of all applicable action sequences; a fact
%% $p$ is a landmark \cite{hoffmann:etal:jair-04} iff $\true$ entails
%% $p$-true-at-some-point; other examples presumably exist. From this
%% point of view, we generalize previous concepts to a broader
%% perspective aimed at addressing arbitrary user questions. At the same
%% time, our approach itself can be viewed as an instance of model
%% checking of planning models \cite{clarke:etal:01},\footnote{There has
%%   been little work on this subject; Vaquero et
%%   al.\ \shortcite{vaquero:etal:keq-13} use Petri nets to capture and
%%   check dynamic aspects of planning models in itSIMPLE.}
%% systematically checking all entailments between plan properties. Again
%% the value of our framework lies in its suitability for XAIP (plus
%% computational gains from considering all \props\ dependencies in
%% unison rather than running individual entailment checks).
%
%% Identifying the trade-off between soft goals in an iterative
%% oversubscription planning process relates, in terms of purpose, to the
%% work by Yu et al.\ \cite{yu:etal:jair-17}. A major difference is that
%% Yu et al.\ address conditional temporal problems, a form of
%% conditional temporal plans, instead of PDDL-style compactly specified
%% general planning problems. This more restricted focus allows to
%% leverage previous conflict analysis methods in that area. It remains a
%% question for future work whether such conflict analysis could inspire
%% more effective analysis methods in our framework.

We introduced a framework for plan-space explanation via plan-property
dependencies. We believe that the framework is useful conceptually as
a problem formulation shaping a relevant part of XAIP. Our techniques
for first instantiations of the framework exhibit reasonable
performance in IPC benchmark studies. The computed explanations are
often small and thus potentially feasible for human inspection.



%% more concrete use cases (robotics, pentesting) user studies
%
%% more general properties than goal dependencies
%
%% beyond classical planning
%
%% deeper why questions: An interesting yet challenging question for
%% future work is how one can automatically identify relevant plan
%% properties \props\ to analyze.  ... mention WHY and HOW questions
%% from NASA discussions? ie WHY finding additional properties zooming
%% in on a given dependency, HOW finding a minimal relaxation of
%% enforced props under which a given dependency disappears.

In future work, the effectiveness of these explanations for human
users remains to be evaluated in user studies. Another important
question is how to address deeper ``why'' questions, asking for the
reasons behind an entailment
$\entails{\plans}{\prop}{\propq}$. Possible ideas are to include
additional properties into \props, elucidating the causal chain
between \prop\ and \propq; or to find a minimal relaxation (superset)
of the plan set \plans\ for which \prop\ no longer entails \propq,
thus elucidating the circumstances under which that entailment
holds. Last but not least, of course our framework and algorithms can
and should be extended to richer planning frameworks and plan property
languages.



\joerg{maybe mention this?}

	Comment: I was thinking about the plan verbalization work from Manuela Veloso's group 
	-- e.g. \url{https://www.ijcai.org/Proceedings/16/Papers/127.pdf} 
	-- and realized there may be some interesting connections here. 
	Granted these are different kinds of plans, motion plans versus plans in general, 
	but...  I wonder if an approach like in this paper can be a better alternative 
	to the template approach in the verbalization work. For example, a PDO versus 
	a concrete PDO could help with "specificity" while planning models do have 
	a rich literature of abstractions that can be readily leveraged in this framework. 
	I guess the point is if it is useful automating the templatization using plan properties. 
	Maybe something to think about... 

	More (recent) details here if you are interested:
	\begin{enumerate}
		\item \url{http://www.cs.cmu.edu/~mmv/papers/18gcai-PereraVeloso.pdf}
		\item (data you can use) \url{http://www.cs.cmu.edu/~vdperera/paper/vdperera_thesis.pdf#page=129}
	\end{enumerate}
