\section{Goal-Property Dependencies in OSP}
\label{framework}

%% \joerg{1--1.5 page: previous generic framework, now instantiated to
%%   OSP; illustrate defs with goal-fact dependencies in nomystery
%%   example}

%% \joerg{ijcai text snippets:}

We next spell out our framework for plan properties, entailment
relations between them, and the forms of explanations we aim at. We do
so in FDR-based OSP as defined above, but in principle our definitions
are generic and can be instantiated for arbitrary plan properties and
planning frameworks. We will discuss the definitions from that
perspective.



\subsection{Plan Properties and Property Entailment}

The plan properties we consider here are formulas over the soft goals:

\begin{definition}[Plan Properties]
\label{def:osp-plan-properties}
Let $\task =
(\vars,\allowbreak\acts,\allowbreak\cost,\allowbreak\init,\allowbreak\goalhard,\allowbreak\goalsoft,\allowbreak\costbound)$
be an OSP task, and \plans\ its set of plans.

A \defined{plan property} is a function $\prop_\phi : \plans \mapsto
\{\true, \false\}$ where $\phi$ is a propositional formula over the
atoms \goalsoft, and $\prop_\phi(\plan) = \true$ iff $\phi$ evaluates
to \true\ under the truth value assignment where $g \in \goalsoft$ is
\true\ iff $g \in \init\apply{\plan}$.
%
$\prop_\phi$ is \defined{conjunctive} if $\phi$ has the form
$\bigwedge_{a \in A} a\allowbreak$ or $\neg \bigwedge_{b \in B} b$.
\end{definition}

Our current analyses consider conjunctive plan properties only. We
will identify $\prop_\phi$ with the characterizing formula $\phi$.

In general, a plan property can be any function mapping a task and an
action sequence to a Boolean value. Examples are temporal plan
trajectory constraints, action-set properties (formulas over subsets
of actions touched/not touched by the plan), deadlines, bounds on
resource consumption, etc. To the extent that such more general
properties can be compiled into goal facts, conjunctive plan
properties can be used to analyze their dependencies. We will explore
this possibility here with a compilation for action-set properties.

We formalize plan-property dependencies as entailment in the space of
plans \plans:

\begin{definition}[\plans-Entailment]
\label{def:pi-entailment}
Let $\task =
(\vars,\allowbreak\acts,\allowbreak\cost,\allowbreak\init,\allowbreak\goalhard,\allowbreak\goalsoft,\allowbreak\costbound)$
be an OSP task, and \plans\ its set of plans.

We say that $\plan \in \plans$ \defined{satisfies} a plan property
$\phi$, written $\plan \models \phi$, if $\prop_\phi(\plan) =
\true$. We denote by $\modelsof{\plans}{\phi} := \{\plan \mid \plan
\in \plans, \plan \models \phi\}$ the subset of plans that satisfy
$\phi$.
%
%% We say that \phi\ is \defined{satisfiable} in \plans\ if
%% $\modelsof{\plans}{\phi} \neq \emptyset$.
%
We say that $\phi$ \defined{\plans-entails} $\psi$, written
$\entails{\plans}{\phi}{\psi}$, if $\modelsof{\plans}{\phi} \subseteq
\modelsof{\plans}{\psi}$.
%
\end{definition}

This definition views $\plans$ in the role traditionally taken by a
knowledge base, identifying a set of ``possible worlds'' within which
entailment over plan properties is considered.
%
Observe that, given this, \plans-entailment is more than standard
entailment: $\phi \Rightarrow \psi$ implies that
$\entails{\plans}{\phi}{\psi}$, but not vice versa. \plans-entailment
captures entailments specific to the space of plans \plans. For
example, in our illustrative NoMystery task, say that all goals are
soft, $T_0$ has initial fuel supply $13$, and $T_1$ has no fuel. Then
$\entails{\plans}{at(P_0,L_4)}{\neg (at(P_1,L_3) \wedge at(P_2,L_5))}$
because, if we achieve the goal for $P_0$, there is insufficient fuel
to transport both other packages. If we set the initial fuel supply of
$T_0$ to $16$, on the other hand, then the knowledge base changes --
\plans\ becomes more permissive -- and that entailment no longer
holds.
%
%$\notentails{\plans}{at(P_0,L_4)}{\neg (at(P_1,L_3) \wedge
%  at(P_2,L_5))}$.

Note that the definition of \plans-entailment is agnostic to the
specification of \plans. The definition applies unchanged to arbitrary
planning frameworks and plan sets \plans. Even an explicitly listed
\plans\ could make sense in some applications.

Our primary focus here will be on goal exclusions:

\begin{definition}[Goal Exclusion]
Let $\task =
(\vars,\allowbreak\acts,\allowbreak\cost,\allowbreak\init,\allowbreak\goalhard,\allowbreak\goalsoft,\allowbreak\costbound)$
be an OSP task, and \plans\ its set of plans.

A \defined{goal exclusion} is an entailment of the form
$\entails{\plans}{\bigwedge_{a \in A} a}{\neg \bigwedge_{b \in B}
  b}$. 
%
The entailment is \defined{non-dominated} if there is no pair
$(A',B')$ where $A' \subseteq A$, $B' \subseteq B$, $(A',B') \neq
(A,B)$, and $\entails{\plans}{\bigwedge_{a \in A'} a}{\neg
  \bigwedge_{b \in B'} b}$.
%
The entailment is \defined{non-rhs-dominated} if there is no $B'$
where $B' \subsetneq B$ and $\entails{\plans}{\bigwedge_{a \in A}
  a}{\neg \bigwedge_{b \in B'} b}$.
\end{definition}

Goal exclusions are of interest in OSP as they reflect the detailed
(soft-)goal trade-offs in the user's quest for a good plan. A
non-dominated goal exclusion has subset-minimal $A$ and $B$. This
dominates entailments with larger $A$ and/or $B$ as it has a weaker
left-hand side $A$ (smaller conjunction) entailing a stronger
right-hand side $\neg B$ (smaller disjunction, after moving the
negation inside). If $A$ is fixed, then only the right-hand side $B$
needs be minimal. We will use non-dominated entailments to give more
compact explanations.

% Note: non-dominated goal exclusions have disjoint A,B: if g in A cap
% B, then A ==> B \ {g} is still valid as g is true whenever A is.


\subsection{Local and Global Explanations}

As previously hinted, we propose to employ the concept of
plan-property entailment for the purpose of giving local and global
explanations of the plan space \plans. 

For local explanations, we assume a user question of the form ``Why do
you not achieve property $\phi$?'', which we answer with the set of
plan properties $\psi$ entailed by $\phi$:

\begin{definition}[Local Explanation (LE)]
\label{def:local-explanation}
Let $\task =
(\vars,\allowbreak\acts,\allowbreak\cost,\allowbreak\init,\allowbreak\goalhard,\allowbreak\goalsoft,\allowbreak\costbound)$
be an OSP task, and \plans\ its set of plans.

For a plan property $\phi$, the \defined{local explanation (LE)} for
$\phi$ is the set $\{\psi \mid \entails{\plans}{\phi}{\psi}\}$ of plan
properties \plans-entailed by $\phi$.
%
For $\phi = \bigwedge_{a \in A} a$, the \defined{goal-exclusion LE}
for $\phi$ is $\{\psi \mid \psi = \neg \bigwedge_{b \in B} b,
\entails{\plans}{\phi}{\psi} \text{ is non-rhs-dominated}\}$.
\end{definition}

Such an answer makes sense if the entailed properties $\psi$ are
undesirable. This is the case, in particular, for goal-exclusion
LEs. In our example with $T_0$ fuel $13$, the answer to $\phi =
at(P_0,L_4)$ ``Why do you not achieve the goal for $P_0$?'' would be
$\psi = \neg (at(P_1,L_3) \wedge at(P_2,L_5))$ ``Because that would
necessitate to forego the goal for either $P_1$ or $P_2$''.

From a general perspective, plan properties here serve as an
abstraction level at which to explain \plans\ to a user. The
underlying assumption is that \plans\ is large and/or the mechanisms
that generate \plans\ are complex, so that an abstract form of
explanation is needed. The abstraction level can be controlled through
the number and granularity of plan properties. Given this, while here
we simply talk about all formulas over soft-goal facts, it can make
sense to instead fix a more specific set \props\ of plan properties
the user has a vested interest in (raising the new sub-problem how to
choose \props).

Note that, if the user question ``Why do you not achieve property
$\phi$?''  refers to a given plan candidate \plan, then a simpler
variant of local explanation is to return only those entailed $\psi$
where $\plan \not\models \psi$, \ie, currently false plan properties
that would become true when enforcing $\phi$. This is easier to
compute and yields smaller explanations. In our experiments, we will
only use explanations as per Definition~\ref{def:local-explanation},
yielding an upper bound on computational hardness and explanation
size, and avoiding the bias incurred by a particular method for
generating candidate plans \plan.

A canonical notion of global explanation arises directly from the
above.  Instead of showing the implications of one specific plan
property, one can show all such implications:

\begin{definition}[Global Explanation (GE)]
Let $\task =
(\vars,\allowbreak\acts,\allowbreak\cost,\allowbreak\init,\allowbreak\goalhard,\allowbreak\goalsoft,\allowbreak\costbound)$
be an OSP task, and \plans\ its set of plans.

Say that $\phi$ and $\psi$ are \defined{\plans-equivalent} if
$\modelsof{\plans}{\phi} = \modelsof{\plans}{\psi}$, and denote by
$\equiv{\plans}{\phi}$ the \plans-equivalence class of $\phi$.
%
The \defined{global explanation (GE)} for \task\ is the partial order
$\pdo{\plans}$ over the classes $\equiv{\plans}{\phi}$ where
$\equiv{\plans}{\phi} \pdo{\plans} \equiv{\plans}{\psi}$ iff
$\entails{\plans}{\phi}{\psi}$.
%
%% A \defined{concrete GE (cGE)} replaces each equivalence class
%% $\equiv{\plans}{\prop}$ with exactly one $\prop \in
%% \equiv{\plans}{\prop}$.
\end{definition}

The only somewhat non-obvious design decision here is to group plan
properties into equivalence classes. That said, while this definition
makes sense at the formal level, it is of doubtful practical
value. There can be many plan properties (here: all propositional
formulas over \goalsoft) and many equivalence classes (here: in the
worst case, one class for every set of truth-value assignments to
$\goalsoft$). This makes it questionable whether the GE can be
processed by a user.
%
% Joerg: de-emphasizing computational aspectb her as not much better
% in restricted def
%
%% has several practicality issues. From a computational perspective,
%% there can be many plan properties (here: all propositional formulas
%% over \goalsoft) and many equivalence classes (here: subsets of
%% satisfying $\plan \in \plans$). Deciding even a single entailment
%% relation can be hard (here: \pspace-complete as it subsumes plan
%% existence \cite{bylander:ai-94}).  the size of a GE makes it more than
%% questionable whether the GE can be processed by a human user.

It therefore makes sense to focus on more limited forms of GEs, and to
find ways to represent these more compactly. Here, we focus on
non-dominated goal exclusions:

\begin{definition}[Goal-Exclusion GE]
Let $\task =
(\vars,\allowbreak\acts,\allowbreak\cost,\allowbreak\init,\allowbreak\goalhard,\allowbreak\goalsoft,\allowbreak\costbound)$
be an OSP task, and \plans\ its set of plans.

The \defined{goal-exclusion GE} for \task\ is the partial order
$\pdo{\plans}$ over conjunctive plan properties where $\phi
\pdo{\plans} \psi$ iff $\entails{\plans}{\phi}{\psi}$ is a
non-dominated goal exclusion.
%
%% $\phi = \bigwedge_{a \in A} a$, $\psi = \neg \bigwedge_{b \in B}
%% b$, $\entails{\plans}{\bigwedge_{a \in A} a}{\neg \bigwedge_{b \in
%% B} b}$, and there does not exist $(A',B')$ such that $A' \subseteq
%% A$, $B' \subseteq B$, $(A',B') \neq (A,B)$, and
%% $\entails{\plans}{\bigwedge_{a \in A'} a}{\neg \bigwedge_{b \in B'}
%% b}$.
\end{definition}

Given the form $\entails{\plans}{\bigwedge_{a \in A} a}{\neg
  \bigwedge_{b \in B} b}$ of goal exclusions, when ignoring all other
\plans-entailments, plan properties can no longer entail each other
and the equivalence classes trivialize.
%
% Joerg: now introduced above
%
%% In words, (1) we consider only entailments where achieving a goal
%% conjunction $A$ necessitates to forego at least one element of a goal
%% conjunction $B$; and (2) we show only those entailments where $A$ and
%% $B$ are subset-minimal. For (1), the motivation is that users are
%% likely to be particularly interested in this form of dependency, as it
%% reflects the detailed (soft-)goal trade-offs in the user's quest for a
%% good plan. Thanks to the exclusive focus on (1), plan properties can
%% no longer entail each other, so that the equivalence classes
%% trivialize. For (2), entailments $\entails{\plans}{\bigwedge_{a \in A}
%%   a}{\neg \bigwedge_{b \in B} b}$ with minimal $A$ and $B$ dominate
%% non-minimal ones, with a weaker left-hand side $A$ (smaller
%% conjunction) entailing a stronger right-hand side $\neg B$ (smaller
%% disjunction, after moving the negation inside). 
%
% Joerg: just leave out, probably better to not ``wake sleeping
% hounds'', should get away with compactness being a good thing
%
%% We assume here that, in an actual user interface, this form of
%% domination is natural to, and cognitively easy for, human users when
%% processing the goal-exclusion GE; this remains to be verified through
%% user studies in future work. For now, o
%
% Joerg: now introduced above
%
Our results on benchmarks suggest that the goal-exclusion GE can be
feasible to compute, and often produces answers sufficiently compact
for human inspection.

Consider again our example, with $T_0$ fuel $13$. It is not difficult
to determine that this example has exactly three non-dominated goal
exclusions, namely $\entails{\plans}{P_0}{\neg (P_1 \wedge P_2)}$,
$\entails{\plans}{P_1}{\neg (P_0 \wedge P_2)}$, and
$\entails{\plans}{P_2}{\neg (P_0 \wedge P_1)}$ when taking the freedom
to notate goal facts by the respective packages.
%
%% $\entails{\plans}{at(P_0,L_4)}{\neg (at(P_1,L_3) \wedge
%%   at(P_2,L_5))}$; $\entails{\plans}{at(P_1,L_3)}{\neg (at(P_0,L_4)
%%   \wedge at(P_2,L_5))}$; and $\entails{\plans}{at(P_2,L_5))}{\neg
%%   (at(P_0,L_4 \wedge at(P_1,L_3))}$.
%
So the size of the goal-exclusion GE, in terms of the ordering
relations it contains, is three. 

Note that this GE is equivalent to the statement ``\goalsoft\ is not
solvable as a whole, but each of its subsets is''. Indeed, as we see
next, the special case of goal-exclusion dependencies is equivalent to
the analysis of minimal unsolvable goal subsets.  This connection
further facilitates compact representation, and we exploit it in our
algorithms.
