\documentclass[12pt]{article}

\input{packeges.tex}

\input{commands.tex}


\begin{document}

\section{Plan-Property Dependency 1}

\textit{If goal subset A is true at the end of the plan, then at least 
one element of goal subset B must be false at the end of the plan.}\\

\noindent
This is equivalent to unsolvability of the goal 
$A \cup B$, i.e., $\bigwedge_{p \in A \cup B} p$.
\noindent
Subsumption relation over such dependencies: simply set inclusion: 
if $A \cup B$ is unsolvable, then any superset is, too.

\paragraph{Task} 
Compute a compact representation of all subsumption-minimal dependencies 
of this kind.

\paragraph{Algorithmic solution to this task} 
Given the simple form of the subsumption relation here, the compact 
representation sought is simply the set of all minimal unsolvable goal 
subsets. This can be computed through a search tree that starts at the 
node $N_0$ containing the complete goal G, and where each search step 
on a node N 

\begin{enumerate}
	\item tests solvability of $N$
	\item  stops if $N$ is solvable
	\item otherwise generates one child node $N'$ for every subset of $N$
	 obtained by removing one fact. 

\end{enumerate}

\noindent
Upon termination, the \textit{unsolvable leaves} of this tree are the minimal 
unsolvable goal subsets.

\textbf{unsolvable leaves:} Unsolvable nodes with only solvable children.

%\textcolor{red}{B always size 1}

\section{Plan-Property Dependency 2}

\textit{If goal subset A is true at the end of the plan, 
then ALL elements of goal subset B must be false at the end of the plan.}\\

\noindent
This is equivalent to unsolvability of the goal 
$\bigwedge_{p \in A} p \wedge (\bigvee_{q \in B} q)$.
Subsumption relation over such dependencies: 
if $\bigwedge_{p \in A} p \wedge (\bigvee_{q \in B} q)$ is unsolvable 
then the same is true for any superset of A and subset of B.

\paragraph{Task}
Compute a compact representation of all subsumption-minimal dependencies of this kind.\\

\noindent
\textcolor{red}{\textbf{Can be derived from PPD1!}}


\newpage
\section{Examples}

Nomystery domain with one truck 6 locations and 5 packages. 
Driving from one location to an other cost one unit of fuel. 
The fuel is scaled from F (minimum fuel to deliver all packages) to 1. \\

\input{example_1}
%\newpage
\input{example_2}


\newpage
\paragraph{PPD 2}
Derivation for example 2 with fuel $ = 5$:

\begin{center}
	PPD1 $ =\{ \{P_0, P_1\} \{P_0, P_2\} \{P_1, P_3\} \{P_0, P_3, P_4\} \{P_2, P_3, P_4\} \}$
\end{center}

\noindent
From PDD1 we get the not necessary optimal (A is minimal, B could be larger) PPD2s in the left table. 

\begin{itemize}
	\item $P \in PDD1$ and $ p \in P$
	\item $A = P \setminus p$
	\item $B = \{p\}$
\end{itemize}

\noindent
To derive the optimal PPD2 for an A, you have to merge all B's which belong 
to an A' which is a subset of A. 

\begin{align*}
	(A, \bigcup_{(A', B') \in \text{ASub}((A, B))} B_j \cup B)
\end{align*}

\noindent
$\text{ASub}((A, B)) = \{(A', B') \in \text{PPD2}| A' \subseteq A\}$.\\

\noindent
If the planning task is not solvable for a goal fact at all, you can add 
this goal fact to all B's.\\

\begin{minipage}[t]{0.35\textwidth}
	\centering
	original PPD2\\
	\begin{tabular}{r|l|l}
		& A & B \\\hline
		1 & $P_0$ & $P_1$ \\
		2 & $P_1$ & $P_0$ \\
		3 & $P_0$ & $P_2$ \\
		4 & $P_2$ & $P_0$ \\
		5 & $P_1$ & $P_3$ \\
		6 & $P_3$ & $P_1$ \\
		7 & $P_0, P_3$ & $P_4$ \\
		8 & $P_0, P_4$ & $P_3$ \\
		9 & $P_3, P_4$ & $P_0$ \\
		10 & $P_2, P_3$ & $P_4$ \\
		11 & $P_2, P_4$ & $P_3$ \\
		12 & $P_3, P_4$ & $P_2$ \\
	\end{tabular}
\end{minipage}
\hfill
\begin{minipage}[t]{0.55\textwidth}
	\centering
	derived PPD2\\
	\begin{tabular}{l|l|l|l}
		& ASub & A & B \\\hline
		1 & 3 & $P_0$ & $P_1, P_2$ \\
		2 & 5 & $P_1$ & $P_0, P_3$ \\
		4 & & $P_2$ & $P_0$ \\
		6 & & $P_3$ & $P_1$ \\
		7 & 1, 3 & $P_0, P_3$ & $P_1, P_2, P_4$ \\
		8 & 1, 3 & $P_0, P_4$ & $P_1, P_2, P_3$ \\
		10 & 4 ,6 & $P_2, P_3$ & $P_0, P_1, P_4$ \\
		11 & 4 & $P_2, P_4$ & $P_0, P_3$ \\
		12 & 6, 9 & $P_3, P_4$ & $P_0, P_1, P_2$ \\	
	\end{tabular}
\end{minipage}

\paragraph{Formal Definition}
Let PPD1 be all minimal unsolvable goal subsets.

\begin{align*}
	PPD2 := \{(A,B) | 
		\exists P \in PPD1:(
			A \subseteq P \wedge |P \setminus A | = 1 \wedge
			\forall P' \in PPD1:
				A \subseteq P' \rightarrow P' \setminus A \subseteq B
		)
	\}
\end{align*}

\newpage
\section{Empirical Performance}

	\subsection{Nomystery}
		\input{experiments_nomystery}

		\newpage
	\subsection{IPC Benchmarks}
		\input{experiments_ipc}

\end{document}